Entrainment of low Mach number thermals in stratified domains

### 2019-05-29 tex/stratified_thermals.tex
small mechanical or clarification changes. Renames some parameterization constants.

%%%%% Body of the paper
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% INTRODUCTION
\section{Introduction}
\label{sec:intro}
Recent observations of solar convection have revealed a convective conundrum.
Power spectra of horizontal velocities show weaker flows than anticipated at large length scales \citep{hanasoge&all2012, greer&all2015}.
These obserations cast doubt on the existence of ``giant cells'' driven by deep convection which should manifest as powerful, large-scale motions at the solar surface and throughout the solar convection zone. 
This discrepancy between theory and observations has called into question our fundamental understanding of convection, sparking numerous targeted investigations the nature of convection in the Sun  \citep{featherstone&hindman2016, omara&all2016, cossette&rast2016, kapyla&all2017, hotta2017}.

@@

If entropy rain does evolve into downward propagating, dense vortex rings, it is important to understand how the filling factor of these basic convective elements is affected by their entropy signature. 
In the context of Earth's atmosphere, ``thermals,'' or buoyant fluid regions which evolve into rising vortex rings, are thought to be the nucleus of cloud formation. 
Atmospheric cloud-forming thermals always rise, but the term is also used for the reverse, falling process.
The evolution of thermals in the Boussinesq limit has been well studied in the laboratory for decades \citep[see e.g.][]{morton&all1956, scorer1957}, and more recently have been studied through Direct Numerical Simulation (DNS) in the laminar and turbulent regime \citep{lecoanet&jeevanjee2018}. 
One fundamental result of these studies of thermals is that they experience a large degree of entrainment characterized by radial expansion and decceleration of propagation velocity.
However, we do not know of a study in which the propagation of these thermals, and thus the nature of their entrainment, is affected by a significant atmospheric stratification.

@@

\subsection{Evolution of momentum and impulse}
The evolution of thermals as buoyant vortex rings has been well described in the unstratified, Boussinesq limit for decades (see e.g. \LJ\,for a description and references).
While many descriptions of vortex rings assume that they evolve in a self-similar state, we will lay out a description of the momentum and impulse of the thermal and show how these quantities control the evolution of the vortex ring.

@@

In the low-Mach number limit in which changes in density from the background are negligible, and in the limit of a thin-core vortex ring, the impulse of a vortex ring can be approximated as
\begin{equation}
I_z \approx \pi \rho r^2 \Gamma,
\label{eqn:impulse_approx}
\end{equation}
where $r$ is the radius of the thermal from its axis of symmetry to the maxima of its buoyant signature, and $\Gamma = \int_{\mathcal{A}} (\grad\times\bm{u})dA$ is the circulation in a cross-section of the vortex ring.
Integrating the momentum equation in the absence of viscous effects shows that changes in circulation in an axisymmetric vortex ring are
\begin{equation}
\frac{d\Gamma}{dt} = \oint_{\mathcal{C}} g \frac{S_1}{c_P}\hat{z} \cdot d\bm{x},
\label{eqn:circulation}
\end{equation}
where $\mathcal{C}$ is the contour around the area $\mathcal{A}$ in which the circulation is contained and $d\bm{x}$ is the line integral element around that contour.
For the case of a vortex core in which the entropic signature is contained tightly in the core, as in Fig.~\ref{fig:evolution_colormeshes}, any contour which is drawn around the vortex core should conserve circulation in the absence of viscosity.
As a result, we will treat the circulation of a developed vortex ring as a conserved quantity.

@@

After spinning up, we assume that the thermal is characterized by a constant buoyant signature, $B_0$, and constant circulation, $\Gamma_0$.
While these thermals began as initial spherical perturbations, they can be modeled as having always been vortex rings which evolved from a ``virtual origin'' where the vortex ring is characterized by having a radius of zero. 
From this virtual origin, the vortex ring's momentum and impulse grow to a value of $M_0$ and $I_0$ at time $t = 0$ when the true thermal is released from rest.

We first note, as in Fig.~\ref{fig:constants}a, that the buoyancy in our thermals is not necessarily perfectly constant in time, and that effects like detrainment result in a minor reduction of the buoyant signature of our thermals. 
We thus express the buoyancy as 
\begin{equation}
B \approx \chi B_0,
\end{equation}
where $\chi$ is a constant of $\mathcal{O}$(1) which represents this detrainment. 
We then integrate Eqn.~\ref{eqn:change_in_impulse} in time,
\begin{equation*}
I_z = \chi B_0 t + I_0.
\end{equation*}

Combining this result with Eqn.~\ref{eqn:impulse_approx}, we retrieve our first result,
\begin{equation}
r = \sqrt{\frac{\chi B_0 t + I_0}{\pi\rho\Gamma_0}}.
\label{eqn:r_theory}
\end{equation}
Following Eqn.~\ref{eqn:circulation}, and taking $\Gamma_0$ to be constant (which is a decent assumption in our numerical experiments, as in Fig.~\ref{fig:constants}b), we find that the radius grows with time but this growth is damped by increasing atmospheric density.
In the Boussinesq limit where $\rho \rightarrow \text{constant}$, we retrieve the $r \propto \sqrt{t}$ scaling found in the Boussinesq regime in \LJ. 
We find that the inclusion of stratification adds the additional complexity of $r \propto \rho^{-1/2}$, such that downward-propagating vortex rings (as studied here) will entrain less than boussinesq thermals, and upward-propagating rings will entrain more. 
In the absence of buoyancy ($B = 0$), this result aligns with the prediction for purely horizontal compression noted by \citet{brandenburg2016} of $r^2 \propto \rho^{-1}$.

The momentum can likewise be simply integrated,
\begin{equation*}
M_z = \beta\chi B_0 t + M_0.
\end{equation*}
For the proper choice of vertical velocity, $w_{\text{th}}$, and volume, $\mathcal{V}$, the momentum is also precisely 
\begin{equation*}
M_z = \rho \mathcal{V} w_{\text{th}}.
\end{equation*}
We approximate the volume as $\mathcal{V} = f r^3$, where $f$ is a parameter which we take to be constant. 
This assumption is perhaps not as perfect as the assumptions of constant $B_0$ or $\Gamma_0$ (see Fig.~\ref{fig:constants}c), but it is nevertheless not a bad assumption. 
Combining our approximate expressions, and using our theoretical description of $r$ (Eqn.~\ref{eqn:r_theory}), we retrieve
\begin{equation}
\rho^{-1/2} w_{\text{th}} = \left(\frac{(\pi \Gamma)^{3/2}}{f}\right)\frac{\beta\chi B_0 t + M_0}{(\chi B_0 t + I_0)^{3/2}}.
\end{equation}
Defining the thermal velocity $w_{\text{th}} \equiv dz_{\text{th}}/dt$, and making the assumption that the vortex ring is at the virtual origin at $t = 0$ so that $I_0 = M_0 = 0$, an integrable expression for the evolution of thermal position over time can be retrieved,
\begin{equation}
\frac{dz_{\text{th}}}{\rho(z_{\text{th}})^{1/2}} =
\left(\frac{\beta(\pi\Gamma)^{3/2}}{f(\chi B_0)^{1/2}}\right)\frac{dt}{t^{1/2}}
\label{eqn:dz_theory}
\end{equation}
If the atmospheric stratification in which the thermal is falling is known, this result can be integrated with the known profile of $\rho(z_{\text{th}})$ in order to find the position of the thermal as a function of time.

@@

\subsection{Solution for thermal evolution in a Polytrope}
In this work, we study an ideal gas whose equation of state is $P = \rho T$ and whose stratification is a perfectly adiabatic polytrope,
\begin{gather}
T_0 = 1 + (\grad_{\text{ad}})(z - L_z), \\
\rho_0 = T_0^{\,m_{\text{ad}}},
\label{eqn:polytrope}
\end{gather}
where $m_{\text{ad}} = (\gamma-1)^{-1}$, and where the adiabatic temperature gradient in our nondimensional systems (see section \ref{sec:experiment}) is $\grad_{\text{ad}} \equiv g(e^{n_\rho/m_{\text{ad}}} - 1)/(L_z c_P)$, with $g = m_{\text{ad}} + 1$ and $L_z = 20$ is the nondimensional depth of the atmosphere in units of thermal diameters.

Integrating Eqn.~\ref{eqn:dz_theory} under this polytropic density stratification and assuming $m_{\text{ad}} < 2$, we find
\begin{equation}
z_{\text{th}} = \grad_{\text{ad}}^{-1}\left[\left(\frac{2C}{ \alpha } \sqrt{t + t_{\text{off}}} + T_0^{1/\alpha}  \right)^{\alpha} - 1\right] + L_z,
\label{eqn:theory_z}
\end{equation}
where $C \equiv \beta \pi^{3/2} \grad_{\text{ad}} / f \sqrt{\Gamma^3/(\chi B)}$ and the temperature at the virtual origin is $T_0 = 1 + \grad_{\text{ad}}(z_0 - L_z)$.
The virtual origin is located at $z = z_0$ and $t = -t_{\text{off}}$ in simulation units, which are the height and time a which the vortex ring would have had zero radius in the absence of a spin-up phase.
The constant $\alpha^{-1} = 1 - m_{\text{ad}}/2$, and in the limit of large stratification, we find that $z_{\text{th}} \propto t^2$ for our case of $\alpha = 4$.

The thermal is initialized as a uniform sphere of dense fluid, and it quickly spins up into a vortex ring. 
While we do not attempt to model the spin-up phase in this paper, it can be parameterized by the buoyancy $B_0$, circulation $\Gamma_0$, as well as the virtual origin $z_0$, and temporal offset $t_{\text{off}}$. 
Our theory also involves the volumetric aspect ratio of the thermal $f$, the detrainment fraction $\chi$, and the effective buoyancy $\beta$. 
These appear to be constant or only weakly dependent on the stratification for the thermals we have simulated.

@@

\subsection{Anelastic Simulations}
The LBR anelastic equations are \citep{lecoanet&all2014},
\begin{gather}
\DivU = -w \partial_z \ln\rho_0, \\
\begin{split}
\partial_t \bm{u} + \bm{u}\cdot\grad&\bm{u} = \\
- \grad \varpi + S_1\hat{z} &
+ \frac{1}{\rho_0\text{Re}}\left[\grad^2 \bm{u} + \frac{1}{3}\grad(\DivU)\right] 
\end{split}\\
\begin{split}
\partial_t S_1 + \bm{u}\cdot\grad S_1 =& \\
\frac{1}{\text{Re}}\left(\frac{1}{\text{Pr}\rho_0c_P }\right.&[\grad^2 S_1 + \partial_z\ln T_0 \cdot\partial_z S_1]\\
&+ \left.\frac{-(\grad_{\text{ad}})}{\rho_0 T_0}\sigma_{ij}\partial_{x_i}u_j \right),
\end{split}
\end{gather}
where $\lilstressT$ is the viscous stress tensor in units of inverse time.
In our azimuthally symetric domain, we assume that $\partial_\phi = 0$; as the initial conditions of our simulations are at rest and have no azimuthal velocity, $u_\phi$, we explicitly impose that $u_\phi = 0$; therefore $\bm{u} = u_r \hat{r} + w\hat{z}$.

These equations have been nondimensionalized in the same manner as in \citet{lecoanet&jeevanjee2018} such that the length scale is the diameter of the initial thermal perturbation and the velocity scale is the freefall velocity. 
The timescale is thus the freefall crossing time of this unit length. 
These equations are then fully specified in terms of the Reynolds number and Prandtl number,
\begin{equation}
\text{Re} = \frac{ u_{\text{th}} L_{\text{th}}}{\nu}, \qquad
\text{Pr} = \frac{ u_{\text{th}} L_{\text{th}}}{\chi}, \qquad
u_{\text{th}}^2 = \frac{g L_{\text{th}} \Delta s}{c_P},
\end{equation}
where $u_{\text{th}}$ is the freefall velocity, $L_{th}$ is the thermal length scale, and
$\Delta s$ is the magnitude of the specific entropy signature of the thermal.

We choose an atmospheric model in which the dynamic viscosity, $\mu = \rho_0 \nu$, and the thermal conductivity, $\kappa = \rho_0 \chi$, are both uniform and constant in time.
The diffusivities $\nu$ and $\chi$ therefore scale inversely with the density.
As the diffusivities scale with depth, Re is specified at the thermal's initial depth.
All simulations conducted in this work use a value of Re = 600 and $\mbox{\text{Pr} = 1}$.

\subsection{Fully Compressible Simulations}
In order to verify the validity of our 2D Anelastic simulations, we evolve select thermals according to the 3D Navier Stokes equations in a cartesian domain. 
We use the $(T, \ln\rho)$ formulation of the equations in which we have previously studied fully compressible convection at low and high Mach number \citep{lecoanet&all2014, anders&brown2017},
\begin{gather}
\frac{\partial \ln\rho_1}{\partial t} + \epsilon^{-1}\left(\bm{u}\cdot\grad\ln\rho_0 + \DivU\right) = -\bm{u}\cdot\grad\ln\rho_1, \\
\begin{split}
\frac{\partial \bm{u}}{\partial t}  +\grad T_1 + T_1\grad\ln\rho_0 + T_0\grad\ln\rho_1  =\\
- \epsilon T_1\grad\ln\rho_1 + \frac{1}{\rho\text{Re}}\left[\grad^2\bm{u} + \frac{1}{3}\grad(\DivU)\right]
\end{split} \\
\begin{split}
\frac{\partial T_1}{\partial t} + \epsilon^{-1}\left[\bm{u}\cdot\grad T_0 + (\gamma-1)T_0\DivU\right] = \\
-\left[\bm{u}\cdot\grad T_1 + (\gamma-1)T_1\DivU\right] + \frac{1}{\rho c_V\text{Re}}\left[\frac{1}{\text{Pr}}\grad^2 T_1 + \sigma_{ij}\partial_{x_i}u_j\right].
\end{split}
\end{gather}
These equations have been nondimensionalized on the same length and timescales as the anelastic equations, use the same atmospheric profiles and assumptions, and we have explicitly assumed that the background atmosphere ($T_0, \ln\rho_0$) is in hydrostatic and thermal equilibrium in the writing of these equations. 
The new parameter $\epsilon = u_{th}^2$ is the magnitude of entropy perturbations and sets the Mach number of the thermal flows, and we use $\epsilon = 10^{-4}$ in this work.

@@

\subsection{Initial conditions}
To initialize the simulation, we specify a spherical initial specific entropy perturbation,
\begin{equation}
S_1 = - \frac{1}{2}\left[1 - \text{erf}\left(\frac{r' - r_{\text{th}}}{\delta}\right)\right].
\label{eqn:thermal_IC}
\end{equation}
Here, $r' = \sqrt{r^2 + (z - z_0)^2}$, where $z_0 = L_z - 3r_{\text{th}}$, with the thermal radius set as $r_{\text{th}} = 0.5$, and a smoothing width, $\delta = 0.1$.
As mentioned previously, Re = 600 and Pr = 1 are specified at the thermal's initial depth, $z = z_0$.

@@

In Fig.~\ref{fig:diff}a, we display measured thermal radius vs. depth, with a detailed comparison of our 2D Anelastic and 3D Fully Compressible cases. 
In the low stratification limit, the radius of the thermal grows linearly with depth, $r \propto d$, aligning with the Boussinesq limit shown in \citet{lecoanet&jeevanjee2018}.
This growth of the thermal is the result of entrainment of environmental fluid and accompanies decceleration like $w \propto d^{-1}$ in the Boussinesq limit, as is shown in Fig.~\ref{fig:results_panels}b.
However, as stratification increases, the thermal entrainment of environmental fluid decreases and it experiences less expansion.
In the limit of large stratification ($n_\rho \geq 3$), thermals begin contracting with depth.
This lessened entrainment is, unsurprisingly, accompanied by greater acceleration of the thermal (Fig.~\ref{fig:results_panels}).

In Fig.~\ref{fig:diff}b, the fractional difference in radius vs. depth between 3D FC and 2D AN cases is shown for the $n_\rho = [0.5, 1, 2]$ cases studied here. 
Aside from early in the simulation, when the thermal is spinning up, and late in the simulation, when the 3D simulations begin interacting with the impenetrable boundary at the bottom of the domain, the 2D anelastic and 3D fully compressible simulations agree surprisingly well.
We find differences in the radius of less than 1\% throughout the bulk of the thermal evolution for the cases studied here.
This gives us confidence that our high stratification anelastic simulations are producing reliable results.
Furthermore, this close agreement between low Mach number Fully Compressible simulations and Anelastic simulations parallels the agreement between the equation sets seen in \citet{lecoanet&all2014}.

The measured values of each of the parameters described in section \ref{sec:theory} for each of our simulations  \ref{sec:theory} for each of our simulations  \ref{sec:theory} for each of our simulations are presented in table \ref{table:parameters}.
As anticipated, $\beta$ is $\mathcal{O}$(0.5) and $\chi$ is $\mathcal{0}$(1), with both decreasing slightly in value with increasing stratification, which is perhaps unsurprising in light of Fig.~\ref{fig:constants}.
In all cases, the buoyancy $B_0$ is similar to the integrated buoyancy in the initial conditions, with some losses due to detrainment in the spin-up.
In our nondimensional units, we also find that the circulation is $\mathcal{0}$(-2) for each of our cases, and decreases with increasing stratification (and decreasing vortex size).

>>>>

%%%%% Body of the paper
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% INTRODUCTION
\section{Introduction}
\label{sec:intro}
Recent observations of solar convection have revealed a convective conundrum.
Power spectra of horizontal velocities show weaker flows than anticipated at large length scales \citep{hanasoge&all2012, greer&all2015}.
These observations cast doubt on the existence of ``giant cells'' driven by deep convection which should manifest as powerful, large-scale motions at the solar surface and throughout the solar convection zone. 
This discrepancy between theory and observations has called into question our fundamental understanding of convection, sparking numerous targeted investigations the nature of convection in the Sun  \citep{featherstone&hindman2016, omara&all2016, cossette&rast2016, kapyla&all2017, hotta2017}.

@@

If entropy rain does evolve into downward propagating, dense vortex rings, it is important to understand how the filling factor of these basic convective elements is affected by their negative buoyancy. 
In the context of Earth's atmosphere, ``thermals,'' or buoyant fluid regions which evolve into rising vortex rings, are thought to be the nucleus of cloud formation. 
Atmospheric cloud-forming thermals always rise, but the term is also used for the reverse, falling process.
The evolution of thermals in the Boussinesq limit has been well studied in the laboratory for decades \citep[see e.g.][]{morton&all1956, scorer1957}, and more recently have been studied through Direct Numerical Simulation (DNS) in the laminar and turbulent regime \citep{lecoanet&jeevanjee2018}. 
One fundamental result of these studies of thermals is that they experience a large degree of entrainment characterized by radial expansion and decceleration of propagation velocity.
However, we do not know of a study in which the propagation of these thermals, and thus the nature of their entrainment, is affected by a significant atmospheric stratification.

@@

\subsection{Evolution of momentum and impulse}
The evolution of thermals as buoyant vortex rings has been well described in the unstratified, Boussinesq limit for decades (see e.g.,~\LJ\,for a description and references).
While many descriptions of vortex rings assume that they evolve in a self-similar state, we will lay out a description of the momentum and impulse of the thermal and show how these quantities control the evolution of the vortex ring.

@@

In the low-Mach number limit in which changes in density from the background are negligible, and in the limit of a thin-core vortex ring, the impulse of a vortex ring can be approximated as
\begin{equation}
I_z \approx \pi \rho r^2 \Gamma,
\label{eqn:impulse_approx}
\end{equation}
where $r$ is the radius of the thermal from its axis of symmetry to the maxima of its buoyant signature, and $\Gamma = \int_{\mathcal{A}} (\grad\times\bm{u})dA$ is the circulation in a cross-section of the vortex ring.
Integrating Eqn.~\ref{eqn:euler_momentum} reveals that changes in circulation can be calculated as
\begin{equation}
\frac{d\Gamma}{dt} = \oint_{\mathcal{C}} g \frac{S_1}{c_P}\hat{z} \cdot d\bm{x},
\label{eqn:circulation}
\end{equation}
where $\mathcal{C}$ is the contour around the area $\mathcal{A}$ in which the circulation is contained and $d\bm{x}$ is the line integral element around that contour.
For the case of a vortex core in which the entropic signature is contained tightly in the core, as in Fig.~\ref{fig:evolution_colormeshes}, any contour which is drawn around the vortex core should conserve circulation in the absence of viscosity.
As a result, we will treat the circulation of a developed vortex ring as a conserved quantity.

@@

After spinning up, we assume that the thermal is characterized by a constant buoyant signature, $B_{\text{th}}$, and constant circulation, $\Gamma_{\text{th}}$.
While these thermals began as initial spherical perturbations, they can be modeled as having always been vortex rings which evolved from a ``virtual origin'' where the vortex ring is characterized by having a radius of zero. 
From this virtual origin, the vortex ring's momentum and impulse grow to a value of $M_0$ and $I_0$ at time $t = 0$ when the true thermal is released from rest.

We first note, as in Fig.~\ref{fig:constants}a, that the buoyancy in our thermals is not necessarily perfectly constant in time, and that effects like detrainment result in a minor reduction of the buoyant signature of our thermals. 
We thus express the buoyancy as 
\begin{equation}
B \approx \chi B_{\text{th}},
\end{equation}
where $\chi$ is a constant of $\mathcal{O}$(1) which represents this detrainment. 
We then integrate Eqn.~\ref{eqn:change_in_impulse} in time,
\begin{equation*}
I_z = \chi B_{\text{th}} t + I_0.
\end{equation*}

Combining this result with Eqn.~\ref{eqn:impulse_approx}, we retrieve our first result,
\begin{equation}
r = \sqrt{\frac{\chi B_{\text{th}} t + I_0}{\pi\rho\Gamma_{\text{th}}}}.
\label{eqn:r_theory}
\end{equation}
Following Eqn.~\ref{eqn:circulation} and taking $\Gamma_{\text{th}}$ to be constant (which is a decent assumption in our numerical experiments, as in Fig.~\ref{fig:constants}b), we find that the radius grows with time but this growth is damped by increasing atmospheric density.
In the Boussinesq limit where $\rho \rightarrow \text{constant}$, we retrieve the $r \propto \sqrt{t}$ scaling found in the Boussinesq regime in \LJ. 
We find that the inclusion of stratification adds the additional complexity of $r \propto \rho^{-1/2}$, such that downward-propagating vortex rings (as studied here) will entrain less than boussinesq thermals, and upward-propagating rings will entrain more. 
In the absence of buoyancy ($B = 0$), this result aligns with the prediction for purely horizontal compression noted by \citet{brandenburg2016} of $r^2 \propto \rho^{-1}$.

The momentum can likewise be simply integrated,
\begin{equation*}
M_z = \beta\chi B_{\text{th}} t + M_0.
\end{equation*}
For the proper choice of vertical velocity, $w_{\text{th}}$, and volume, $\mathcal{V}$, the momentum is also precisely 
\begin{equation*}
M_z = \rho \mathcal{V} w_{\text{th}}.
\end{equation*}
We approximate the volume as $\mathcal{V} = f r^3$, where $f$ is a parameter which we take to be constant. 
This assumption is perhaps not as perfect as the assumptions of constant $B_{\text{th}}$ or $\Gamma_{\text{th}}$ (see Fig.~\ref{fig:constants}c), but it is nevertheless not a bad assumption. 
Combining our approximate expressions, and using our theoretical description of $r$ (Eqn.~\ref{eqn:r_theory}), we retrieve
\begin{equation}
\rho^{-1/2} w_{\text{th}} = \left(\frac{(\pi \Gamma_{\text{th}})^{3/2}}{f}\right)\frac{\beta\chi B_{\text{th}} t + M_0}{(\chi B_{\text{th}} t + I_0)^{3/2}}.
\end{equation}
Defining the thermal velocity $w_{\text{th}} \equiv dz_{\text{th}}/dt$, and making the assumption that the vortex ring is at the virtual origin at $t = 0$ so that $I_0 = M_0 = 0$, an integrable expression for the evolution of thermal position over time can be retrieved,
\begin{equation}
\frac{dz_{\text{th}}}{\rho(z_{\text{th}})^{1/2}} =
\left(\frac{\beta(\pi\Gamma_{\text{th}})^{3/2}}{f(\chi B_{\text{th}})^{1/2}}\right)\frac{dt}{t^{1/2}}.
\label{eqn:dz_theory}
\end{equation}
If the atmospheric stratification in which the thermal is falling is known, this result can be integrated with the known profile of $\rho(z_{\text{th}})$ in order to find the position of the thermal as a function of time.

@@

\subsection{Solution for thermal evolution in a Polytrope}
In this work, we study an ideal gas with an adiabatic index of $\gamma = 5/3$, whose equation of state is $P = \rho T$, and whose stratification is a perfectly adiabatic polytrope,
\begin{gather}
T_0 = 1 + (\grad_{\text{ad}})(z - L_z), \\
\rho_0 = T_0^{\,m_{\text{ad}}},
\label{eqn:polytrope}
\end{gather}
where $m_{\text{ad}} = (\gamma-1)^{-1}$, and where the adiabatic temperature gradient in our nondimensional systems (see section \ref{sec:experiment}) is $\grad_{\text{ad}} \equiv g(e^{n_\rho/m_{\text{ad}}} - 1)/(L_z c_P)$, with $g = m_{\text{ad}} + 1$ and $L_z = 20$ is the nondimensional depth of the atmosphere in units of thermal diameters.
All thermodynamic quantities are nondimensionalized such that $\rho_0 = T_0 = 1$ at $z = L_z$.

Integrating Eqn.~\ref{eqn:dz_theory} under this polytropic density stratification and assuming $m_{\text{ad}} < 2$, we find
\begin{equation}
z_{\text{th}} = \grad_{\text{ad}}^{-1}\left[\left(\frac{2C}{ \alpha } \sqrt{t + t_{\text{off}}} + T_{\text{th},0}^{1/\alpha}  \right)^{\alpha} - 1\right] + L_z,
\label{eqn:theory_z}
\end{equation}
where $C \equiv \beta \pi^{3/2} \grad_{\text{ad}} / f \sqrt{\Gamma_{\text{th}}^3/(\chi B_{\text{th}})}$ and the temperature at the virtual origin is $T_{\text{th},0} = 1 + \grad_{\text{ad}}(z_{\text{th},0} - L_z)$.
The virtual origin is located at $z = z_0$ and $t = -t_{\text{off}}$ in simulation units, which are the height and time at which the vortex ring would have had zero radius in the absence of a spin-up phase.
The constant $\alpha^{-1} = 1 - m_{\text{ad}}/2$, and in the limit of large stratification, we find that $z_{\text{th}} \propto t^2$ for our case of $\alpha = 4$.

The thermal is initialized as a uniform sphere of dense fluid, and it quickly spins up into a vortex ring. 
While we do not attempt to model the spin-up phase in this paper, it can be parameterized by the buoyancy $B_{\text{th}}$, circulation $\Gamma_{\text{th}}$, as well as the virtual origin $z_{\text{th,0}}$, and temporal offset $t_{\text{off}}$. 
Our theory also involves the volumetric aspect ratio of the thermal $f$, the detrainment fraction $\chi$, and the effective buoyancy $\beta$. 
These appear to be constant or only weakly dependent on the stratification for the thermals we have simulated.

@@

\subsection{Anelastic Simulations}
The LBR anelastic equations are \citep{lecoanet&all2014},
\begin{gather}
\DivU = -w \partial_z \ln\rho_0, \\
\begin{split}
\partial_t \bm{u} + \bm{u}\cdot\grad&\bm{u} = \\
- \grad \varpi + S_1\hat{z} &
+ \frac{1}{\rho_0\text{Re}}\left[\grad^2 \bm{u} + \frac{1}{3}\grad(\DivU)\right] 
\end{split}\\
\begin{split}
\partial_t S_1 + \bm{u}\cdot\grad S_1 =& \\
\frac{1}{\text{Re}}\left(\frac{1}{\text{Pr}\rho_0c_P }\right.&[\grad^2 S_1 + \partial_z\ln T_0 \cdot\partial_z S_1]\\
&+ \left.\frac{-(\grad_{\text{ad}})}{\rho_0 T_0}\sigma_{ij}\partial_{x_i}u_j \right),
\end{split}
\end{gather}
where $\lilstressT$ is the viscous stress tensor in units of inverse time.
In our azimuthally symetric, cylindrical domain, we assume that $\partial_\phi = 0$; as the initial conditions of our simulations are at rest and have no azimuthal velocity, $u_\phi$, we explicitly impose that $u_\phi = 0$; therefore $\bm{u} = u_r \hat{r} + w\hat{z}$.

These equations have been nondimensionalized in the same manner as in \LJ\, such that the length scale is the diameter of the initial thermal perturbation and the velocity scale is the freefall velocity. 
The timescale is thus the freefall crossing time of this unit length. 
These equations are then fully specified in terms of the Reynolds number and Prandtl number,
\begin{equation}
\text{Re} = \frac{ u_{\text{th}} L_{\text{th}}}{\nu}, \qquad
\text{Pr} = \frac{ u_{\text{th}} L_{\text{th}}}{\chi}, \qquad
u_{\text{th}}^2 = \frac{g L_{\text{th}} \Delta s}{c_P},
\end{equation}
where $u_{\text{th}}$ is the freefall velocity, $L_{th}$ is the thermal length scale, and
$\Delta s$ is the magnitude of the specific entropy signature of the thermal.

We choose an atmospheric model in which the dynamic viscosity, $\mu = \rho_0 \nu$, and the thermal conductivity, $\kappa = \rho_0 \chi$, are both uniform and constant in time.
We make this choice as the mass-weighted diffusivities, $\mu$ and $\kappa$, appear in the density-weighted momentum and entropy equations, and we find the density-weighted entropy and momentum to be key quantities in our thermal theory.
The diffusivities $\nu$ and $\chi$ therefore scale inversely with the density.
As the diffusivities scale with depth, Re is specified at the thermal's initial depth.
All simulations conducted in this work use a value of Re = 600 and $\mbox{\text{Pr} = 1}$.

\subsection{Fully Compressible Simulations}
In order to verify the validity of our 2D Anelastic simulations, we evolve select thermals according to the 3D Navier Stokes equations in a cartesian domain. 
We use the $(T, \ln\rho)$ formulation of the equations in which we have previously studied fully compressible convection at low and high Mach number \citep{lecoanet&all2014, anders&brown2017},
\begin{gather}
\frac{\partial \ln\rho_1}{\partial t} + \epsilon^{-1}\left(\bm{u}\cdot\grad\ln\rho_0 + \DivU\right) = -\bm{u}\cdot\grad\ln\rho_1, \\
\begin{split}
\frac{\partial \bm{u}}{\partial t}  +\grad T_1 + T_1\grad\ln\rho_0 + T_0\grad\ln\rho_1  =\\
- \epsilon T_1\grad\ln\rho_1 + \frac{1}{\rho\text{Re}}\left[\grad^2\bm{u} + \frac{1}{3}\grad(\DivU)\right]
\end{split} \\
\begin{split}
\frac{\partial T_1}{\partial t} + \epsilon^{-1}\left[\bm{u}\cdot\grad T_0 + (\gamma-1)T_0\DivU\right] = \\
-\left[\bm{u}\cdot\grad T_1 + (\gamma-1)T_1\DivU\right] + \frac{1}{\rho c_V\text{Re}}\left[\frac{1}{\text{Pr}}\grad^2 T_1 + \sigma_{ij}\partial_{x_i}u_j\right].
\end{split}
\end{gather}
These equations have been nondimensionalized on the same length and time scales as the anelastic equations, use the same atmospheric profiles and assumptions, and we have explicitly assumed that the background atmosphere ($T_0, \ln\rho_0$) is in hydrostatic and thermal equilibrium in the writing of these equations. 
The new parameter $\epsilon = u_{\text{th}}^2$ is the magnitude of entropy perturbations and sets the Mach number of the thermal flows, and we use $\epsilon = 10^{-4}$ in this work.

@@

\subsection{Initial conditions}
To initialize the simulation, we specify a spherical initial specific entropy perturbation,
\begin{equation}
S_1 = \frac{1}{2}\left[\text{erf}\left(\frac{r' - r_{\text{th}}}{\delta}\right) - 1\right].
\label{eqn:thermal_IC}
\end{equation}
Here, $r' = \sqrt{r^2 + (z - z_0)^2}$, where $z_0 = L_z - 3r_{\text{th}}$, with the thermal radius set as $r_{\text{th}} = 0.5$, and a smoothing width, $\delta = 0.1$.
As mentioned previously, Re = 600 and Pr = 1 are specified at the thermal's initial depth, $z = z_0$.

@@

In Fig.~\ref{fig:diff}a, we display measured thermal radius vs. depth, with a point-by-point comparison of our 2D Anelastic and 3D Fully Compressible cases. 
In the low stratification limit, the radius of the thermal grows linearly with depth, $r \propto d$, aligning with the Boussinesq limit shown in \LJ.
This growth of the thermal is the result of entrainment of environmental fluid and accompanies decceleration like $w \propto d^{-1}$ in the Boussinesq limit, as is shown in Fig.~\ref{fig:results_panels}b.
However, as stratification increases, the thermal entrainment of environmental fluid decreases and it experiences less expansion.
In the limit of large stratification ($n_\rho \geq 3$), thermals begin contracting with depth.
This lessened entrainment is, unsurprisingly, accompanied by greater acceleration of the thermal (Fig.~\ref{fig:results_panels}).

In Fig.~\ref{fig:diff}b, the fractional difference in radius vs. depth between 3D fully compressible and 2D anelastic cases is shown for the $n_\rho = [0.5, 1, 2]$ cases studied here. 
Aside from early in the simulation, when the thermal is spinning up, and late in the simulation, when the 3D simulations begin interacting with the impenetrable boundary at the bottom of the domain, the 2D anelastic and 3D fully compressible simulations agree well.
We find differences in the radius of less than 1\% throughout the bulk of the thermal evolution for the cases studied here.
This gives us confidence that our high stratification anelastic simulations are producing reliable results.
Furthermore, this close agreement between low Mach number Fully Compressible simulations and Anelastic simulations parallels the agreement between the equation sets seen in \citet{lecoanet&all2014}.

The measured values of each of the parameters described in section \ref{sec:theory} for each of our simulations are presented in table \ref{table:parameters}.
As anticipated, $\beta$ is $\mathcal{O}$(0.5) and $\chi$ is $\mathcal{O}$(1), with both decreasing slightly in value with increasing stratification, which is perhaps unsurprising in light of Fig.~\ref{fig:constants}.
In all cases, the buoyancy $B_{\text{th}}$ is similar to the integrated buoyancy in the initial conditions, with some losses due to detrainment in the spin-up.
In our nondimensional units, we also find that the circulation is $\mathcal{O}$(-2) for each of our cases, and decreases with increasing stratification (and decreasing vortex size).
